---
title: "Lecture Notes - 9/10/19"
date: 09/10/2019
output: html_notebook
---

```{r}
fact = 1
print(paste0("1! is ", fact))
```

Rewrite this as a `for loop`
```{r}
fact = 1
for (i in 1:10) {
  fact = i * fact
  print(paste0(i, "! is ", fact))
}
```

|Header1|Header2|
|---|---|
|trying out a table|with data|

## Conditionals 
```{r}
a = c(2,5,6,88,3,20)

if(a[3] < 10){
  print("yes")
}else{
  print("no")
}
```

```{r}
for (i in 1:length(a)){
  if(a[i] < 10){
    print("yes")
  }else{
    print("no")
  }
}
```

# Descriptive Stats

```{r}
setwd("~/Google Drive/Grad School/Fall 2019/BTRY_6010")
data = read.table("titanic(4).txt",header=TRUE)
```

## Histograms
- like a bar graph but the x-axis has numerical meaning
- When first looking at a dataset, histograms are a good place to start
- Clustering
  - No Bumps - uniform
  - 1 bump - unimodal
  - 2 bumb - bimodal
- *Center* where is the middle
- *Spread* how much variablitity
- *Skewness* lack of symmetry
- *Outliers* extreme values that stand out


```{r}
dat = read.table("/Users/mdarfler/Google Drive/Grad School/Fall 2019/BTRY_6010/titanic(4).txt",header=TRUE)
head(dat)

hist(dat$Age)
```

## Measures of central tendancy
- mean
  - $$\bar{x} = \dfrac{1}{n} \sum_{n=1}^{n} x$$
- meadian
- mode

```{r}
x = c(2,1,2,8,2)
mean(x)

sort(x)
```


```{r}
age = dat$Age[!is.na(dat$Age)]
length(age)

age_sort = sort(age)
age_sort[length(age_sort)/2] #median
```

##Sample Median, Quartiles and Percentiles

- given _n_ values x1 ... xn

Median - Equal area
Mean = balancing point

### Weighted Average

If a discrete random variable tekes the values $x_{1}, ..., x_{n}$ whith frequences $f_{1}...f_{k}$ it's average can be calculated as
-
$$\dfrac{x_{1}f_{1} + x_{2}f_{2} + ... + x_{n}f_{n}}{f_{1} + f_{2} + ... +f_{n}}$$

# Measure of Spread or Dispersion
