---
title: "Simple Random Sampling to estimate one population proportion"
author: "Sumanta Basu"
date: "September 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "~/Google Drive/Grad School/Fall 2019/BTRY_6010")
baseurl = "https://raw.githubusercontent.com/mdarfler/BTRY_6010/master/"
```

## Background

The status of $155$ enrolled students in a BTRY 6010 class from a previous semester are listed in enrollment.csv. Using simple random samples of different sizes, we will estimate the proportion of PhD students (population parameter $p$) in class (code 'GR' in the file). The other student status codes are 0C0 (professional development), 1P1 - 3P1 (masters) and 41 (undergraduate senior).

The goal is to demonstrate how simple random sampling helps control the margin of error in our estimation. In particular, we will see that with a sample of size $n$, the proportion of PhD students in our sample ($\hat{p}$) is roughly within $\pm 1/\sqrt{n}$ of the population proportion of PhD students ($p$) (i.e. the proportion of PhD students among these $155$ students). In the process, we will also see some examples of using `if` and `for` statements in R. 

Note that in this case the population ($155$ students in class) is observed, so we can easily calculate the true proportion. However, the point of the exercise is to illustrate how we can get reasonably close to $p$ without using data on all the students, but only a small number ($n$) of them.

We will use the following informal result (we will see a more formal treatment of this result in textbook section 6.1):

``Suppose we select a simple random sample of $n$ students in the class, and find that $X$ of them are PhD students. Then the sample proportion $\hat{p} = X/n$ is a good estimator of $p$, in the sense that there is at least a $95$\% chance the random interval $(\hat{p} - \frac{1}{\sqrt{n}}, \hat{p} + \frac{1}{\sqrt{n}})$ contains $p$.''



## Reading Enrollment Data

```{r}
dat = read.csv(paste0(baseurl,'Lab2/enrollment(3).csv'), header=TRUE, as.is=TRUE)
pop = dat$Level
table(pop)
```

## Population proportion

The true proportion of PhD students in class (which we will estimate using data on smaller samples) is
```{r}
p = mean(pop == 'GR')
p
sum(pop=="GR")/length(pop)
```

## Set random number seed for reproducibility
```{r}
set.seed(51)
```


## Estimate p using a SRS of size 15
```{r}
sample1 = sample(pop, 15, replace=FALSE)
sample1
p.hat = mean(sample1 == 'GR')
p.hat
```
Construct an approx. 95% confidence interval
```{r}
ci.lower = p.hat - (1/sqrt(15))
ci.upper = p.hat + (1/sqrt(15))
print(c(ci.lower, ci.upper))
```
Does this interval contain the true proportion of PhD students in class?

Let's try again. Repeat using another SRS of size 15
```{r}
sample2 = sample(pop, 15, replace=FALSE)
sample2
p.hat = mean(sample2 == 'GR')
p.hat
```
Construct an approx. 95% confidence interval
```{r}
ci.lower = p.hat - (1/sqrt(15))
ci.upper = p.hat + (1/sqrt(15))
print(c(ci.lower, ci.upper))
```
How about now? Did we capture the true proportion?

Let's try this a hundred times, and see how often our interval captures the true proportion
```{r}
n=15 # sample size
success = rep(0, 1000)
for (iter in 1:1000){
  paste0('iter=', iter)
  smpl = sample(pop, n, replace=FALSE)
  p.hat = mean(smpl == 'GR')
  if (((p.hat - (1/sqrt(n))) < p) & (p.hat + (1/(sqrt(n))) > p))
    success[iter] = 1
}
mean(success)
```
## Reducing margin of error with larger sample size
With a sample size of 15, our margin of error is
```{r}
n=15
1/sqrt(n)
```
A 25% margin of error may be considered high. What if we increase our sample size to 50? Our margin of error will now be
```{r}
n = 50
1/sqrt(n)
```
This is narrower margin of error. But does it preserve our level of confidence about the estimate? To check this, let's repeat our procedure a hundred times with n=50, and see how often our interval captures the true proportion
```{r, echo=TRUE}
n=50 # sample size
numTrials = 1000
success = rep(0, numTrials)
for (iter in 1:numTrials){
  paste0('iter=', iter)
  smpl = sample(pop, n, replace=FALSE)
  p.hat = mean(smpl == 'GR')
  if (((p.hat - (1/sqrt(n))) < p) & (p.hat + (1/(sqrt(n))) > p))
    success[iter] = 1
}
mean(success)
```




